---
title: "emp_final_code"
author: "Hy Pham"
date: "5/25/2020"
output:
  word_document: default
  html_document: default
---
```{r}
setwd("~/OneDrive - The CIty College of New York/mbp/m.a. econ/empirical research/emp proj/data")
load("chs_2016.RData")
```

```{r message= FALSE, warning=FALSE}
library(lmtest)
library(sandwich)
library(AER)
library(tidyverse)
library(ggplot2)
library(devtools)
library(dotwhisker)
options(scipen=999)
```

```{r message= FALSE, warning=FALSE}
library(officer)
library(stargazer)
library(flextable)
```

depression
```{r}
summary(chs_2016$currdepress)
summary(chs_2016$phq8score) #patient health queestionnaire depression scale 0-24
```

substances
```{r}
summary(chs_2016$smoker)     #1=never 2=current 3 = former
summary(chs_2016$heavydrink16)
```

demographics. cannot access genderid since gotta get permission so not straight is used
```{r}
summary(chs_2016$age18_64) # 1 = 18024 2=25-44 3=45-64
summary(chs_2016$sex) #1=male
summary(chs_2016$sexualid16) #1=gay/les 2=straight 3 = bi 3=sth else
summary(chs_2016$education) #1=less hs 2=hs 3=some college4=college
summary(chs_2016$emp3) #1=emp 2= unemp 3 =notin laborp
summary(chs_2016$imputed_pov200) #1 = <200
summary(chs_2016$newrace) #1= white 2=black 3=hispanic 4=azn 5=other
summary(chs_2016$hhsize) 
```

converting and filtering out NA's

```{r}
nyc_chs <- select(chs_2016,currdepress, phq8score, smoker, heavydrink16, age18_64, sex, sexualid16, education, emp3, imputed_pov200, newrace, hhsize)
is.na(nyc_chs) <- nyc_chs == ".d"
is.na(nyc_chs) <- nyc_chs == ".r"
is.na(nyc_chs) <- nyc_chs == "."
nyc_chs <- na.omit(nyc_chs)
```

#variables as factors for brief summary

factors
```{r}
sex_factor <- as.factor(nyc_chs$sex)
levels(sex_factor) <- c("Male", "Female")
sex_id_factor <- as.factor(nyc_chs$sexualid16)
levels(sex_id_factor) <- c("Gay/Lesbian", "Straight", "Bi-Sexual", "Something Else")
race_factor <- as.factor(nyc_chs$newrace)
levels(race_factor) <- c("White", "Black", "Hispanic", "Asian", "Other")
education_factor <- as.factor(nyc_chs$education)
levels(education_factor) <-c("Less than HS", "High School Graduate", "Some College", "College Graduate")
```

```{r}
summary(sex_factor)
summary(race_factor)
```


converting to dummies

depression dummies. depressed = yes to depression and phq score >= 10
```{r}
nyc_chs$Depression <- as.numeric(nyc_chs$currdepress == "1")
nyc_chs$Depression_Score <- as.numeric(nyc_chs$phq8score >= "10")
```
siummary
```{r}
summary(nyc_chs$phq8score)
summary(nyc_chs$currdepress)
```

substance dummies
```{r}
nyc_chs$Current_Smoker <- as.numeric(nyc_chs$smoker == "1")
nyc_chs$Heavy_Drinker <- as.numeric(nyc_chs$heavydrink16 == "1")
```

demographics

genders
```{r}
nyc_chs$Female <- as.numeric(nyc_chs$sex == "2")
nyc_chs$Male <- as.numeric(nyc_chs$sex == "1")
nyc_chs$Straight <- as.numeric(nyc_chs$sexualid16 == "2")
nyc_chs$Gay_Lesbian_Bi_Other <- as.numeric(nyc_chs$sexualid16 == "1" | nyc_chs$sexualid16 == "3" | nyc_chs$sexualid16 == "4")
```


age

```{r}
nyc_chs$Age18_24 <- as.numeric(nyc_chs$age18_64 == "1")
nyc_chs$Age25_44 <- as.numeric(nyc_chs$age18_64 == "2")
nyc_chs$Age45_64 <- as.numeric(nyc_chs$age18_64 == "3")
```
race
```{r}
nyc_chs$White <- as.numeric(nyc_chs$newrace == "1")
nyc_chs$Black <- as.numeric(nyc_chs$newrace == "2")
nyc_chs$Hispanic <- as.numeric(nyc_chs$newrace =="3")
nyc_chs$Asian <- as.numeric(nyc_chs$newrace == "4")
nyc_chs$OtherRaces <- as.numeric(nyc_chs$newrace == "5")
```

education
```{r}
nyc_chs$Less_HS<-as.numeric(nyc_chs$education=="1")
nyc_chs$HS_Grad<-as.numeric(nyc_chs$education=="2")
nyc_chs$Some_College<-as.numeric(nyc_chs$education=="3")
nyc_chs$College_Grad<-as.numeric(nyc_chs$education=="4")
```

employment & income & household
```{r}
nyc_chs$Employed <- as.numeric(nyc_chs$emp3 == "1")
nyc_chs$Unemployed <- as.numeric(nyc_chs$emp3 == "2")
nyc_chs$Poor_HH <- as.numeric(nyc_chs$imputed_pov200 == "1")
nyc_chs$Not_Poor_HH <- as.numeric(nyc_chs$imputed_pov200 =="2")
nyc_chs$Average_Household_Size <- as.numeric(mean(nyc_chs$hhsize))
```



phd8 graph
```{r}
phq_chart <- ggplot(nyc_chs, aes(phq8score))
phq_chart + geom_histogram(fill="#FF9999", colour = "black") + ggtitle("The Eight-Item Patient Health Questionnaire Scores Among Responders")+xlab("PHQ-8 Scores\n(Greater than 10 = Depressed)") + ylab("Frequency\n(n=6476)") + coord_flip() +theme(plot.title=element_text(face = "bold",hjust=0.5))
```



sexualid chart
```{r}
par(mfrow=c(1,2))
sexid_chart <- ggplot(nyc_chs,aes(sex_id_factor))
sexid_chart + geom_bar(width=.75, fill="#56B4E9", colour ="black") + ggtitle("Responses to Questions of Sexual Identification") + xlab("Sexual Identity") + ylab("Frequency\n(n=6476)") +theme(plot.title=element_text(face = "bold",hjust=0.5)) + geom_text(stat='count',aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.25)

race_chart <- ggplot(nyc_chs, aes(race_factor))
race_chart + geom_bar(width=0.5, fill = "#009E73", colour = "black") + ggtitle("Population Division by Race/Ethnicity") + xlab("Ethnicity") +ylab("Frequency\n(n=6476)") +theme(plot.title=element_text(face = "bold",hjust=0.5))  + geom_text(stat='count',aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.25)
```

education chart
```{r}
edu_chart <- ggplot(nyc_chs, aes(education_factor))
edu_chart + geom_bar(width = .5, fill ="#0072B2", colour = "black") + ggtitle("Population Division by Education Level") +xlab("Education Level") + ylab("Frequency\n(n=6476)") + theme(plot.title=element_text(face = "bold", hjust =0.5)) + geom_text(stat='count',aes(label=..count..), position=position_dodge(width=0.9), vjust=-0.25)



```




looking for sums

dep_score and race and gender
```{r}
sum(nyc_chs$Depression_Score & nyc_chs$White)
sum(nyc_chs$Depression_Score&nyc_chs$Asian)
sum(nyc_chs$Depressio_Score &nyc_chs$Hispanic)
sum(nyc_chs$Depression_Score & nyc_chs$OtherRaces)
sum(nyc_chs$Depression_Score & nyc_chs$Black)

sum(nyc_chs$Depression_Score & nyc_chs$Straight)
sum(nyc_chs$Depression_Score & nyc_chs$Gay_Lesbian_Bi_Other)
sum(nyc_chs$Depression_Score & nyc_chs$Male)
sum(nyc_chs$Depression_Score & nyc_chs$Female)
```


depression and race and gender
```{r}
sum(nyc_chs$Depression & nyc_chs$White)
sum(nyc_chs$Depression&nyc_chs$Asian)
sum(nyc_chs$Depression &nyc_chs$Hispanic)
sum(nyc_chs$Depression & nyc_chs$OtherRaces)
sum(nyc_chs$Depression & nyc_chs$Black
    )
sum(nyc_chs$Depression & nyc_chs$Straight)
sum(nyc_chs$Depression & nyc_chs$Gay_Lesbian_Bi_Other)
sum(nyc_chs$Depression & nyc_chs$Male)
sum(nyc_chs$Depression & nyc_chs$Female)

```



depression numbers are smaller than the phq scale numbers ->> look at percentages

percentage of phq per race and gender
```{r}
#white 
(886/1816)*100
#asian
(340/850)*100
#hispanic
(1114/2082)*100
#other
(93/172)*100
#black
(714/1556)*100


#straigh
(2919/6085)*100
#other_id
(228/391)*100
#male
(1305/2806)*100
#female
(1842/3616)*100
```

% of depression per race and gender
```{r}
#white 
(128/1816)*100
#asian
(32/850)*100
#hispanic
(260/2082)*100
#other
(21/172)*100
#black
(127/1556)*100


#straigh
(508/6085)*100
#other_id
(60/391)*100
#male
(227/2806)*100
#female
(341/3616)*100
```
percentage phq data frames
```{r}
phq_percentage_race <- data.frame(White = c(48.79),
                                  Asian = c(40.00),
                                  Hispanic = c(53.53),
                                  Other = c(54.07),
                                  Black = c(45.89))
phq_percentage_gender <- data.frame("Straight" = c(47.97),
                                    "Not Straight"=c(58.31),
                                    "Male"=c(46.51),
                                    "Female"=c(50.94))
```




percentage depression data frames
```{r}
dep_percentage_race <- data.frame("White" = c(7.05),
                                  "Asian" = c(3.76),
                                  "Hispanic" = c(12.49),
                                  "Other" = c(12.21),
                                  "Black" = c(8.16))
dep_percentage_gender <- data.frame("Straight" = c(8.35),
                                    "Not Straight"=c(15.35),
                                    "Male"=c(8.09),
                                    "Female"=c(9.43))
```



tables for the above
```{r}
table1 <- flextable(phq_percentage_race)
table1 <- add_header_lines(table1, values = c("Percentage of Respondents with PHQ Greater than 10 Respective to Their Ethnicity"))
table2<-flextable(phq_percentage_gender)
table2 <- add_header_lines(table2, values = c("Percentage of Respondents with PHQ Greater than 10 Respective to Their Gender"))


table3 <-flextable(dep_percentage_race)
table3<-add_header_lines(table3, values=c("Percentages of Responsdents who Said 'Yes' to having Depression in the Last 2 weeks Respective to Their Ethnicity"))
table4<- flextable((dep_percentage_gender))
table4<- add_header_lines(table4, values=c("Percentages of Responsdents who Said 'Yes' to having Depression in the Last 2 weeks Respective to Their Gender"))
print(table1, preview="docx")
print(table2, preview="docx")
print(table3, preview="docx")
print(table4, preview="docx")
```


REGRESSIONS and also regressing 


no control
```{r}
 model11 <- lm(nyc_chs$Depression_Score ~ nyc_chs$Gay_Lesbian_Bi_Other) 
summary(model11)
```

gender model vs phq (male is control)
```{r}
model1 <- lm(nyc_chs$Depression_Score ~ nyc_chs$Gay_Lesbian_Bi_Other  + nyc_chs$Female )
summary(model1)
```

gender model vs dep
```{r}
model2 <- lm(nyc_chs$Depression ~ nyc_chs$Gay_Lesbian_Bi_Other +nyc_chs$Female)
summary(model2)
```


phq vs 
```{r}
model3 <- lm(nyc_chs$Depression_Score ~ nyc_chs$Gay_Lesbian_Bi_Other 
             + nyc_chs$Age25_44 + nyc_chs$Age18_24 
             + nyc_chs$White + nyc_chs$Black + nyc_chs$Hispanic + nyc_chs$Asian
             + nyc_chs$Poor_HH 
             + nyc_chs$Less_HS + nyc_chs$HS_Grad  + nyc_chs$College_Grad + nyc_chs$Unemployed)
summary(model3)
```



```{r}
model4<- lm(nyc_chs$Depression ~ nyc_chs$Gay_Lesbian_Bi_Other 
             + nyc_chs$Age25_44 + nyc_chs$Age18_24 
             + nyc_chs$White + nyc_chs$Black + nyc_chs$Hispanic + nyc_chs$Asian
             + nyc_chs$Poor_HH 
             + nyc_chs$Less_HS + nyc_chs$HS_Grad  + nyc_chs$College_Grad + nyc_chs$Unemployed)
summary(model4)
```


substance
```{r}
model5 <- lm(nyc_chs$Depression_Score~ nyc_chs$Heavy_Drinker + nyc_chs$Current_Smoker + nyc_chs$Poor_HH + nyc_chs$Unemployed)
summary(model5)
```



```{r}
model6 <- lm(nyc_chs$Depression~ nyc_chs$Heavy_Drinker + nyc_chs$Current_Smoker + nyc_chs$Poor_HH + nyc_chs$Unemployed)
summary(model6)
```

```{r}
model_phq <- lm(nyc_chs$Depression_Score ~ nyc_chs$Gay_Lesbian_Bi_Other +nyc_chs$Female
             + nyc_chs$Age25_44 + nyc_chs$Age18_24 
             + nyc_chs$White + nyc_chs$Black + nyc_chs$Hispanic + nyc_chs$Asian
             + nyc_chs$Poor_HH 
             + nyc_chs$Less_HS + nyc_chs$HS_Grad  + nyc_chs$College_Grad + nyc_chs$Unemployed
             + nyc_chs$Heavy_Drinker + nyc_chs$Current_Smoker )
summary(model_phq)
```


```{r}
model_dep <- lm(nyc_chs$Depression ~ nyc_chs$Gay_Lesbian_Bi_Other +nyc_chs$Female
             + nyc_chs$Age25_44 + nyc_chs$Age18_24 
             + nyc_chs$White + nyc_chs$Black + nyc_chs$Hispanic + nyc_chs$Asian
             + nyc_chs$Poor_HH 
             + nyc_chs$Less_HS + nyc_chs$HS_Grad  + nyc_chs$College_Grad + nyc_chs$Unemployed
             + nyc_chs$Heavy_Drinker + nyc_chs$Current_Smoker )
summary(model_dep)
```

```{r}
summary(nyc_chs$Depression_Score)
summary(nyc_chs$Depression)

```

```{r word_table, comment=""}
stargazer(model_phq, model_dep, type = "latex", out="rergession.txt",
          dep.var.labels = c("PHQ-8 Score Greater than 10", "Sel-Reported Depression"),
          covariate.labels = c("Non-Cisgender", "Female",
                               "Age 25-44", "Age 18-24",
                               "White", "Black", "Hispanic", "Asian", 
                               "<200FPL Hoouseholds", 
                               "Did Not Graduate HS", "High School Graduate", "College Graduate",
                               "Heavy Drinking", "Currently Smoking"),
          title="Determinants of Mental Health")
```


```{r}
dwp <- dwplot(list(model_phq, model_dep), show_intercept = TRUE)

dwp + relabel_predictors(c(nyc_chs$Gay_Lesbian_Bi_Other = "Non-Cisgender",
                            nyc_chs$Female = "Female",
                            nyc_chs$Age25_44 = "Age 25-44",
                            nyc_chs$Age18_24 = "Age 18-24",
                            nyc_chs$White = "White",
                            nyc_chs$Black = "Black",
                            nyc_chs$Hispanic = "Hispanic",
                            nyc_chs$Asian = "Asian",
                            nyc_chs$Poor_HH = "<200FPL Households",
                            nyc_chs$Less_HS = "Did Not Graduate HS",
                            nyc_chs$HS_Grad = "High School Graduate",
                            nyc_chs$College_Grad = "College Graduate",
                            nyc_chs$Unemployed = "Unemployment",
                            nyc_chs$Heavy_Drinker = "Heavy Drinking",
                            nyc_chs$Current_Smoker = "Currently Smoking"))+ xlab("Coefficient Estimate") + ylab("") + ggtitle("Determinants of Mental Health")
```

```{r}
dwplot(list(model_phq, model_dep), show_intercept = TRUE) + ggtitle("Determinants of Mental Health") + xlab("Coefficient Estimates")

```








```{r}
dwplot(list(model_phq,model_dep), show_intercept=TRUE, vline =geom_vline(xintercept=0, colour = "grey60", linetype =2)) %>% relabel_predictors(c("nyc_chs$Gay_Lesbian_Bi_Other"= "Non-Cisgender",
                            "nyc_chs$Female" = "Female",
                            "nyc_chs$Age25_44" = "Age 25-44",
                            "nyc_chs$Age18_24" = "Age 18-24",
                            "nyc_chs$White" = "White",
                            "nyc_chs$Black" = "Black",
                            "nyc_chs$Hispanic" = "Hispanic",
                            "nyc_chs$Asian" = "Asian",
                            "nyc_chs$Poor_HH" = "<200FPL Households",
                            "nyc_chs$Less_HS" = "Did Not Graduate HS",
                            "nyc_chs$HS_Grad" = "High School Graduate",
                            "nyc_chs$College_Grad" = "College Graduate",
                            "nyc_chs$Unemployed" = "Unemployment",
                            "nyc_chs$Heavy_Drinker" = "Heavy Drinking",
                            "nyc_chs$Current_Smoker" = "Currently Smoking")) + theme_bw() + xlab("Coefficient Estimate") + ggtitle("Determinants of Mental Health")
```




plot dataframe
```{r}
df1 <- data.frame(white = 48.8, asian = 40)
df1 %>% 
  t() %>% # convert columns to rows
  as.data.frame() %>% # need to change it to a data frame as t() converts it to a matrix
  rownames_to_column("id") %>% 
  ggplot(aes(x = id, y = V1)) +
  geom_col()
```























